---
title: "Figure 5 Identification of aging-related metabolites varying in tissues."
author: "Feng Zhang"
date: "2025/3/18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Clean workspace
rm(list=ls()) 
```

### a. Bar plots displaying the count of differential metabolites between the young and aged tissues.

```{r barplot, echo=TRUE}

library(ggplot2)

ggplot(df,aes(Tissue,value,fill=group))+
  geom_col(width = 0.8)+
  coord_flip()+   
  scale_fill_manual(values = col)+  
  theme_bw()+  
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(color = "black", size = 10),
        axis.text.y = element_text(color = "black",size = 10),        
        axis.title = element_text(color = "black",size = 12),        
        legend.position = "top",
        legend.title = element_blank(),        
        legend.text = element_text(color = "black",size = 12))+  
  labs(x=NULL,y="Aging-related metabolites")+  
  scale_y_continuous(breaks = seq(-200, 200, 100))

```

### b. Aging-related change score of different metabolite classes in each tissue. 

```{r score, echo=TRUE}

library(tidyverse)
library(ggplot2)
library(grid)

x <- max(data$input1)*(-1.15)
y <- max(data$input2)
subtype1.col <- c("#ffe7e1","#73828F","#2c415d","#A7BCBC","#8ba5ae","#fdb762","#a18499","#fdd0a6","#df9175","#b15928","#fff6f0")
col_subt <- c("Alkaloids and derivatives" = subtype1.col[1],"Benzenoids" = subtype1.col[2],"Lipids and lipid-like molecules" = subtype1.col[3],"Nucleosides, nucleotides, and analogues" = subtype1.col[4],"Organic acids and derivatives" = subtype1.col[5],"Organic nitrogen compounds" = subtype1.col[6],"Organic oxygen compounds" = subtype1.col[7],"Organoheterocyclic compounds" = subtype1.col[8],"Organosulfur compounds" = subtype1.col[9],"Others" = subtype1.col[10],"Phenylpropanoids and polyketides" = subtype1.col[11])

g_colors <- c("#513965FF",adjustcolor("#513965FF", alpha.f = 0.6),adjustcolor("#513965FF", alpha.f = 0.1))
h_colors <- c(adjustcolor("#9E4058FF", alpha.f = 0.1),adjustcolor("#9E4058FF", alpha.f = 0.6), "#9E4058FF")
g <- rasterGrob(matrix(g_colors, nrow = 1), width = unit(1, "npc"), height = unit(1, "npc"), interpolate = TRUE)
h <- rasterGrob(matrix(h_colors, nrow = 1),width = unit(1, "npc"), height = unit(1, "npc"), interpolate = TRUE)
ggplot(data) +
  annotation_custom(g, xmin = -Inf, xmax = 0, ymin = -Inf, ymax = Inf) +
  annotation_custom(h, xmin = 0, xmax = Inf, ymin = -Inf, ymax = Inf) +
  geom_point(aes(input2, input1, color = group, size = abs(input2) * 2)) +
  scale_color_manual(values = col_subt) +
  xlab("Age-related Change Score") + 
  ylab("Number of Age-related Metabolites") +  
  theme_bw() + 
  theme(panel.grid = element_blank())+
  theme(legend.position = c(-1.15, 0.2))

```

### e. Inter-tissue correlation heatmaps displaying the metabolic coupling in the young (bottom left) and aged (top right) mice

```{r volcano, echo=TRUE}

# Advanced Heatmap Plots was performed using the OmicStudio tools at https://www.omicstudio.cn.
# Clustering Calculation (hclust): stats package
# Plotting: pheatmap package, version 1.0.12

```

### f. KEGG pathway enrichment analysis of aging-related metabolites in various tissues. 

```{r volcano, echo=TRUE}

library(data.table)
library(reshape2)
library(ggplot2)
library(dplyr)
library(patchwork)

m1 <- data1[data1[,3] %in% data2[1:3,], ] 
m1 <- m1 %>% filter(Sig.Sign.P != "ns") %>% select(c(1, 3, 7, 8))  
m1$log_pvalue <- ifelse(m1$P.value == 0, min(m1$P.value[m1$P.value > 0]), m1$P.value)
p <-ggplot(m1, aes(x=Tissue,y=kegg_pathway_Name)) +
  geom_point(aes(size=Rich.Factor,color=-log10(log_pvalue))) +
  scale_color_gradientn('-log10(Pvalue)',colors=my_palette) + #
  theme_bw() +labs(x=(""),y=(""))+
  scale_y_discrete(position = "right")+
  theme(legend.position="right",
        panel.grid.major.x = element_blank(),
        axis.ticks = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 10, color = "black"),
        panel.border = element_rect(size = 0.7, linetype = "solid", colour = "black"),
        plot.margin = unit(c(0.5,0.5,1,1), "lines"))
p

```

