---
title: "Figure 4 Organ-resolved proteomic change and disease-associated clues."
author: "Feng Zhang"
date: "2025/3/18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Clean workspace
rm(list=ls()) 
```

### c. The top pathways enriched in organ-resolved proteins for each organ

```{r network, echo=TRUE}

## Pathway enrichment analysis of organ-resolved proteins were performed by metascape (https://metascape.org/gp/index.html). Kappa-test score were calculated between each two representative terms selected from the metascape results (P value ≤ 0.05) and set as similarity score between terms. 

## Cytoscape (version 3.9.1) was used for visualization of representative pathway terms, terms were set as nodes and similarity scores that more than 0.3 were set as edges.

```

### e. The number of organ-resolved proteins that are shared by more than one organ

```{r volcano, echo=TRUE}

library(UpSetR)
library(tidyr)
library(ggplot2)

p <-  upset(df2[-1],
      sets= names(df2)[-1], 
      nintersects=50, 
      point.size=1.8, 
      mb.ratio = c(0.55,0.45),
      line.size = 1,
      sets.bar.color = NA,
      main.bar.col= "#3b3b3b",
      matrix.color= "#3b3b3b",
      text.scale = c(1, 1, 1, 1, 1, 1),
      sets.x.label = "",
      mainbar.y.label  = "Intersection size",
      queries = list(
        list(query = elements,params = list("Brain","Lung"),color = "#9E4058FF",active = TRUE),
        list(query = elements,params = list("Heart","Pancrea"),color = "#9E4058FF",active = TRUE),
        list(query = elements,params = list("Cerebellum","Spinal.cord"),color = "#9E4058FF",active = TRUE)))
p


```

### h. Bubble plots displaying organ distributions of aging-upregulated or aging-downregulated organ-resolved proteins

```{r volcano, echo=TRUE}

library(dplyr)
library(ggplot2)
library(ggrepel)
library(cowplot)

xmin <- -0.2
xmax <-  0.2
x_axis_min <- -1.8
x_axis_max <-  1.8
y_axis_min <-  2.2
y_axis_max <-  4.7
xlab <- "log2FC (Proteome)"
ylab <- "Tissue-specific score across tissues"
x_tick_pos <- seq(x_axis_min, x_axis_max, 0.6)
y_tick_pos <- seq(y_axis_min, y_axis_max, 0.5)

data <- data %>% mutate(quadrant = case_when(
  x <= xmin  ~ "Q1",
  x >  xmin & x <= xmax ~ "Q2",
  x >  xmax  ~ "Q3"))
head(data)

annotate_x <- c(rep(x_axis_max, 3), rep((x_axis_min+x_axis_max)/2, 3), rep(x_axis_min, 3))
annotate_y <- rep(c(y_axis_max, (y_axis_min+y_axis_max)/2, y_axis_min), 3)
annotate_text_color <- c(rep("black", 4), "white", rep("black", 4))

ggplot(data) + 
  geom_rect(aes(xmin=-1.8,xmax=-0.2,ymin=-Inf,ymax=Inf),fill="#FFD8DF",alpha=0.3)+
  geom_rect(aes(xmin=0.2,xmax=1.8,ymin=-Inf,ymax=Inf),fill="#C7DFE8",alpha=0.3)+
  geom_point(aes(x=x, y=y, color=group), size=1) +
  coord_cartesian(xlim=c(x_axis_min, x_axis_max),ylim=c(y_axis_min, y_axis_max)) + 
  facet_wrap( ~ group, nrow=3,  scales = "fixed")+
  labs(x=xlab, y=ylab) + 
  theme_bw() + theme(legend.position="none", panel.grid=element_blank()) + 
  scale_colour_manual(values=c("#78AECB", "#639FB0", "#98C7A5","#83C180", "#B2A4A5", "#EC8D63","#CFC397", "#F6B279", "#6197B4","#CEA168", "#A6A952", "#9ACC90","#E97679", "#DE4247", "#A38CBD","#795FA3", "#E0C880", "#C28B65","#A65A34", "#DE4B3F", "#DD9E82")) + 
  scale_x_continuous(breaks=x_tick_pos) +
  scale_y_continuous(breaks=y_tick_pos)+ 
  theme(title = element_text(size = 10,colour = 'black'),
        axis.text.y = element_text(size = 12,colour = 'black'),
        axis.text.x = element_text(size = 12,colour = 'black'),
        axis.title.x = element_text(size = 13,colour = 'black'),
        axis.title.y = element_text(size = 13,colour = 'black'))

```

### i. circular connection diagram displaying organ distributions of aging-upregulated or aging-downregulated organ-resolved proteins

```{r circular plot, echo=TRUE}

library(ggplot2)
library(dplyr)
library(scales)

Glycan$angle <- ifelse(Glycan$Node < 1660,
                       75-(150/1658*(Glycan$Node-1)),
                       75-(150/1658*(Glycan$Node-1))-30)
Glycan$G.x <- cos(Glycan$angle*pi/180.0)*(nrow(protein)/2)
Glycan$G.y <- sin(Glycan$angle*pi/180.0)*(nrow(protein)/2)
head(Glycan)

# Outlier
curve <- Glycan[,3, drop=F]
curve$x.start <- Glycan$G.x*1.1
curve$x.end <- Glycan$G.x*1.15
curve$y.start <- Glycan$G.y*1.1
curve$y.end <- Glycan$G.y*1.15
curve <- curve[-c(276:337,358:419,451:512,565:626,655:716,832:892,917:978,1155:1216,1337:1398,1584:1645,1734:1795,
                  1852:1913,2084:2145,2259:2320,2375:2436,2478:2539,2665:2726,2798:2859,3014:3075,3199:3260),] 
head(curve)

# Central
protein.pos <- protein[,1:2]
head(protein.pos)
protein.pos <- arrange(protein.pos, Count)
protein.pos$P.x <- 0
protein.pos$P.y <- seq(floor(nrow(protein)/2),-floor(nrow(protein)/2))
protein.pos$cluster <- ifelse(protein.pos$Count < 4, protein.pos$Count, ifelse(protein.pos$Count < 6, "4", "5"))
head(protein.pos)

# Connect line
connect <- protein[,c(1,3)]
head(connect)
Gly.name <- strsplit(as.character(connect[,2]),split=";")
connect.count <- sapply(Gly.name,length)
connection <- data.frame(Protein=rep(connect$Protein,connect.count),Glycan=unlist(Gly.name))

#merge
all<-read.table("all.xls",sep = "\t",check.names = F,stringsAsFactors = F,header = T)
head(all)
all1 <- merge(all,protein.pos,by="Protein")
all1$seg.clu <- ifelse(all1$Glycan=="HexNAc(2)Hex(6)Phospho(1)","zz",all1$Gly.clu)
head(all1)

cols <- c("#78AECB", "#639FB0", "#98C7A5","#83C180", "#B2A4A5", "#EC8D63","#CFC397", "#F6B279", "#6197B4",
          "#CEA168", "#A6A952", "#9ACC90","#E97679", "#DE4247", "#A38CBD","#795FA3", "#E0C880", "#C28B65",
          "#A65A34", "#DE4B3F", "#DD9E82")
fills <- c("#FFD8DF","#AFAFAF","#C7DFE8")
p <- ggplot(all1) +   
  geom_point(aes(G.x,G.y,color=Gly.clu))+ 
  geom_segment(aes(x=G.x,y=G.y,xend=P.x,yend=P.y,color=seg.clu),alpha=0.1)+ 
  geom_rect(aes(xmin=-1,xmax=1,ymin=P.y,ymax=P.y+1,fill=cluster))+
  geom_segment(data=curve,aes(x=x.start,xend=x.end,y=y.start,yend=y.end,color=Gly.clu),size=5)+
  theme_minimal() +theme(legend.position = "none",axis.text = element_blank(),
                         axis.title = element_blank(),panel.grid = element_blank())+
  scale_color_manual(values=cols) +
  scale_fill_manual(values=fills)

```

### k. Disease ontology enrichment results in aging-related tissue-resolved proteins of each tissue.

```{r DOSE, echo=TRUE}

library(DOSE)
library(org.Mm.eg.db)
library(clusterProfiler)

uniprot_ids <- c()
data <- bitr(uniprot_ids, fromType = "UNIPROT", toType = "ENTREZID", OrgDb = org.Hs.eg.db) 
do_enrich <- enrichDO(gene = data$ENTREZID,ont = "DO",pAdjustMethod = "BH",qvalueCutoff = 0.05,readable = TRUE)
head(do_enrich)

## Cytoscape (version 3.9.1) was used for visualization of specific terms

```


