---
title: "Figure 1 Characterization of age-related cross-tissue proteome and metabolome in mammals."
author: "Feng Zhang"
date: "2025/3/18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Clean workspace
rm(list=ls()) 
```

### b-c. Principal component analyses (PCA) plot showing proteomic and metabolomic clustering based on tissue or age

```{r pca, echo=TRUE}

# Perform PCA with scaling
pca_result <- prcomp(df, scale. = TRUE)
pca_scores <- as.data.frame(pca_result$x)

# Calculate the proportion of variance explained by each principal component
explained_var <- pca_result$sdev^2 / sum(pca_result$sdev^2)

# Add grouping variable 
pca_scores$group <- factor(c(rep("Group1", n1), rep("Group2", n2)))  

# Create PCA plot
pca_plot <- ggplot(pca_scores, aes(x = PC1, y = PC2, color = group)) +
  geom_point(size = 3) +  # Scatter plot of PCA scores
  stat_ellipse(type = "norm", level = 0.95) +  # Add 95% confidence ellipses
  xlab(paste0("PC1 (", round(explained_var[1] * 100, 2), "%)")) +  # Label x-axis with explained variance
  ylab(paste0("PC2 (", round(explained_var[2] * 100, 2), "%)")) +  # Label y-axis with explained variance
  theme_minimal() +  # Use a minimal theme
  ggtitle("PCA Plot with 95% Confidence Ellipses")  # Add a title
print(pca_plot)

```

### d-e. Relative ratios of the number of detected, differential and detected/differential proteins or metabolites across tissues.

```{r radar, echo=TRUE}

library(ggradar)
library(cols4all)
library(ggplot2)

mycol<-c("#513965FF","#9E4058FF","#BCBD30")

ggradar(dt,
        grid.min = 0, grid.mid = 0.3, grid.max = 1.1,  
        gridline.min.colour = "black", gridline.mid.colour = "black", gridline.max.colour = "black",  
        base.size = 1, axis.label.size = 4, grid.label.size = 6, 
        values.radar = c("", "30%", ""),  
        group.colours = mycol,
        group.point.size = 3, 
        group.line.width = 1, 
        background.circle.transparency = 0,
        legend.position = 'right',
        legend.text.size = 8,
        fill = TRUE,
        fill.alpha = 0.1)

```


### f-g left. Scatterplots of the normalized enrichment scores (NES) for features from proteomic and metabolomic pathway enrichment analyses in the thymus or pancreas

```{r NES, echo=TRUE}

library(tidyverse)
library(ggplot2)

ggplot(data) +
  geom_vline(xintercept = 0, linewidth=1,linetype = "dashed", color = "#DFDFDF") + # 添加垂直虚线
  geom_hline(yintercept = 0, linewidth=1,linetype = "dashed", color = "#DFDFDF") + # 添加水平虚线
  geom_point(aes(input1,input2,color=group),size=7) +
  geom_smooth(aes(input1,input2),span = 2, method='lm', color = "#513965FF") +
  scale_color_manual(values = "#AFAFAF")+
  xlab("NES(Proteomics)") + 
  ylab("NES(Metabolics)") +  
  theme_bw() + 
  theme(panel.grid = element_blank()) + #去除网格线
  theme(legend.position = c(-1.15, 0.2))
ggsave("NES consistency.pdf", width = 5, height = 5)

```

### f-g right. Significant pathway terms enriched in both proteome and metabolome for young or aged thymus or pancreas

```{r Pathway, echo=TRUE}

library(ggstatsplot)
library(data.table)
library(dplyr)
library(tidyr)
library(ggplot2)

pc <- ggplot(data,aes(NES_Pro,forcats::fct_reorder(Pro,NES_Pro))) +
  geom_point(aes(NES_Pro,col=NES_Pro),size=5) +
  geom_point(aes(NES_Met,col=NES_Met),size=5,shape=17) +
  scale_colour_gradientn(colours=c("#513965FF","#EAEBEDFF","#9E4058FF")) +
  scale_size_continuous(range =c(2,8))  +
  scale_y_discrete(guide = guide_axis(position = "right")) +
  theme_bw()+
  theme(axis.text = element_text(size = 10, colour = "black"))+
  theme(axis.title.x = element_text(size = 14, colour = "black"))+
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5, vjust = 1))+ 
  labs(x = "NES(Thymus/Pancreas)", y =NULL)
ggsave(pc, file="Pathway consistency.pdf",height=5,width = 8.3)

```


